---
author: "Jyoti Ankam"
title: "hw_3"
date: "April 4, 2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = F)
knitr::opts_chunk$set(message = F)
```

Loading the library ISLR
```{r}

library(ISLR)

```

Viewing the Weekly dataset
```{r}

data(Weekly)
names(Weekly)
View(Weekly)


```
a) SUMMARIES OF THE DATASET:

Let's use the summary() fuction to see the numerical summaries of the dataset

```{r}

summary(Weekly)

```

Let's see some graphical summaries

```{r}
cor(Weekly[,-9])
plot(Weekly$Volume)
pairs(Weekly)
```

#Summaries

As evident, all the Lags (1 to 5) have similar means and medians which implies that the return percentage has no correlation with time.
Also from the plots above (for each pair), we can see that there is a correlation between Year and Volume - Volume increases as Year increases. Apart of these two variables, no other variables display any kind of relationship. 

b) Using the full data set to perform a logistic regression with Direction as the response and the five Lag variables plus Volume as predictors:

# Logistic Regression
```{r}

log.reg <- glm(Direction ~ Lag1 + Lag2 + Lag3 + Lag4 + Lag5 + Volume, family=binomial, data= Weekly)
summary(log.reg)

```

As seen in the results, only Lag2 is statistically significant with a p-value of 0.0296 which is < 0.05
All the other predictors are statisitcally insignificant. Higher p-values indicates that there is not enough evidence to reject Null hypothesis. With p-value of 0.0296, Lag2 displays some statistical significance.

c) Computing the confusion matrix and overall fraction of correct predictions. Briefly explaining what the confusion matrix is telling you:

# Confusion Matrix

```{r}

log.probs <- predict(log.reg,type="response")
log.probs[1:10]

log.pred <- ifelse(log.probs > 0.5, "Up", "Down")

log.table <- table(log.pred,Weekly$Direction)

mean <- (log.table["Down", "Down"] + log.table["Up", "Up"])/nrow(Weekly)
error <- 1 - mean

View(mean) #0.5610
View(error)#0.4389

```

The % of correct predictions is 56.1% without making Type 1 or 2 errors. This means that we were able to perdict the correct trend 56.1% times and also indicates that our prediction is wrong 43.9 % times. 

A confusion matrix is a table that is often used to describe the performance of a classification model (or "classifier") on a set of test data for which the true values are known. The number at the 


